### 2019-08-04 appendix/Online_appendix_2019.tex
up to the gender preference section, starting appendix work

\documentclass[11pt]{article}

\usepackage{amssymb,amsmath,amsthm}
\usepackage{mathrsfs}
\usepackage{bbm}
\usepackage{apacite}
\usepackage{pstricks, sgamevar, egameps}
\usepackage{chapterbib}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cases}
\usepackage{url}
\usepackage{placeins}
\usepackage{longtable}

\topmargin=-0.45in      
\evensidemargin=0in     
\oddsidemargin=0in      
\textwidth=6.5in        
\textheight=9.0in       
\headsep=0.25in

\usepackage[margin = 1in]{geometry}

\newcommand*{\graphicPath}{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/matlabwages/"}% 
\newcommand*{\TablePath}{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/tables/"}% 
\graphicspath{{\graphicPath}}
\makeatletter
\def\input@path{{\TablePath}}
\makeatother

\title{Online Appendix}
\author{Miriam Larson-Koester}
\date{Fall 2019}

\begin{document}

\maketitle

\newpage

\section{More Model Results} \label{modelresults}
\subsection{Model Job Parameter Estimates by Year}
\input{Estimation_occfinal_year_1960}
\clearpage
\input{Estimation_occfinal_year_1970}
\clearpage
\input{Estimation_occfinal_year_1980}
\clearpage
\input{Estimation_occfinal_year_1990}
\clearpage
\input{Estimation_occfinal_year_2000}
\clearpage
\input{Estimation_occfinal_year_2012}

\section{Lifetime Income Estimation}
\subsection{Comparison of Lifetime Income Estimates to PSID}

%\includegraphics[width=\textwidth]{SIPPlifeinc}
\includegraphics[width=\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPP_lifeinc_occfinal_1960"}
\subsubsection{Histogram of lifetime income: PSID ages 25-55}
\includegraphics[width=\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc"}

\begin{center}
\clearpage
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPPlifeinc_age_occfinal_0"}
\newline
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc_age_occfinal_0"}
\clearpage
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPPlifeinc_age_occfinal_1"}
\newline
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc_age_occfinal_1"}
\clearpage
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPPlifeinc_age_occfinal_2"}
\newline
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc_age_occfinal_2"}
\clearpage
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPPlifeinc_age_occfinal_3"}
\newline
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc_age_occfinal_3"}
\clearpage
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPPlifeinc_age_occfinal_4"}
\newline
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc_age_occfinal_4"}
\clearpage
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPPlifeinc_age_occfinal_5"}
\newline
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc_age_occfinal_5"}
\clearpage
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/SIPPlifeinc_age_occfinal_6"}
\newline
\includegraphics[width=.8\textwidth]{"/Users/Miriam/OneDrive/Box Sync/TYPlocal/output/graphs/PSIDlifeinc_age_occfinal_6"}
\clearpage
\end{center}

\section{Model Fit Wage Histograms}
\subsubsection{Model Fit for Years 2000/2012}

\includegraphics[width=.8\textwidth]{logwages11reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages12reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages13reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages14reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages15reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages16reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages17reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages18reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages19reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages20reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages21reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages22reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages23reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages24reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages25reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages26reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages27reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages28reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages29reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages30reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages31reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages32reg1trunc_occfinal}
\clearpage
\includegraphics[width=.8\textwidth]{logwages33reg1trunc_occfinal}
\newline
\includegraphics[width=.8\textwidth]{logwages34reg1trunc_occfinal}

\end{document}