### 2020-09-22 website stuff/CV Miriam Larson-Koester Fall 2020.tex
fall 2020 cv

% Jason R. Blevins - Curriculum Vitae
%
% Copyright (C) 2004-2013 Jason R. Blevins
%
% You may use use this document as a template to create your own CV
% and you may redistribute the source code freely.  No attribution is
% required in any resulting documents.  I do ask that you please leave
% this notice and the above URL in the source code if you choose to
% redistribute this file.

\documentclass[10pt,letterpaper]{article}

\usepackage{hyperref}
\usepackage{geometry}
\usepackage[shortlabels]{enumitem}
\usepackage{hyperref}

% Fonts
%\usepackage[T1]{fontenc}
%\usepackage[urw-garamond]{mathdesign}

% Set your name here
\def\name{Miriam Larson-Koester}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {economics, gender, labor},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}

\geometry{
  body={6.5in, 9.0in},
  left=1.0in,
  top=1.0in
}

% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\thispagestyle{empty}

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

% Make lists without bullets and compact spacing
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
    \setlength{\itemsep}{0.25em}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0.25em}
  }
}{
  \end{list}
}

\begin{document}

% Place name at left
{\huge \name}

% Alternatively, print name centered and bold:
%\centerline{\huge \bf \name}

\bigskip

\begin{minipage}{0.37\linewidth}
{\bf Contact Information:}\\
Bureau of Economics \\
Federal Trade Commission\\
600 Pennsylvania Ave NW \\
 Washington, DC 20580 
\end{minipage}
\begin{minipage}{0.63\linewidth}
  \begin{tabular}{ll}
    Email: & \href{mailto:mlarsonkoester@ftc.gov}{mlarsonkoester@ftc.gov} \\
    Homepage: & \href{https://sites.google.com/site/miriamlarsonkoester/}{https://sites.google.com/site/miriamlarsonkoester/} \\
    Twitter: & \href{https://twitter.com/KoesterLarson}{@KoesterLarson}
  \end{tabular}
\end{minipage}
\vspace{3.5mm}

\section*{\scshape Current Employment}
\begin{itemize}
 \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
	\item Economist, Federal Trade Commission \hfill{2018-Present}

\end{itemize}

\section*{\scshape Research Interests}
\begin{itemize}
 \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
	\item Labor Economics, Applied Econometrics, Industrial Organization
\end{itemize}

\section*{\scshape Education}

\begin{itemize}
 \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
	\item Ph.D. Economics, Cornell University \hfill{2018}

	\hspace{15mm} \textit{Committee:} Francesca Molinari ({\it chair}), Richard Mansfield, Victoria Prowse
  	\item M.A. Economics, Cornell University \hfill{2015}
	\item B.A. Economics, \textit{Summa Cum Laude}, Mount Holyoke College \hfill{2012}

\end{itemize}

\section*{\scshape Research}
%
%\subsection*{Job Market Paper}
%\begin{itemize}
%\item {\bf A Model of Automobile Supply with Endogenous Product Characteristics: Comparing CAFE Standards and Feebates}
%\end{itemize}
%
%\subsection*{Publications}
%\begin{itemize}
%\item {\bf Is there an Energy Paradox in Fuel Economy? A Note on the Role of Consumer Heterogeneity and Sorting Bias} (with Antonio Bento and Shanjun Li) {\it Economics Letters}
%\end{itemize}

%\subsection*{Publications}
%\begin{itemize}
%
%\item { Low-Income Housing Development, Poverty Concentration, and Neighborhood Inequality} (with Matthew Freedman), {\it Journal of Policy Analysis and Management}, 2015, 34(4): 805-834.
%
%\end{itemize}

\subsection*{Publications}
\begin{itemize}
\item { Is There Still Son Preference in the United States?} (with Francine D. Blau, Lawrence M. Kahn, Peter Brummund, and Jason Cook). Forthcoming at \textit{Journal of Population Economics}. \href{https://link.springer.com/article/10.1007/s00148-019-00760-7?wt_mc=Internal.Event.1.SEM.ArticleAuthorOnlineFirst&utm_source=ArticleAuthorContributingOnlineFirst&utm_medium=email&utm_content=AA_en_06082018&ArticleAuthorContributingOnlineFirst_20200107}{\underline{link}}  \href{http://www.nber.org/papers/w23816}{\underline{NBER}} 
\end{itemize}

\subsection*{Working Papers}
\begin{itemize}
\item { Occupation Gender Segregation: Empirical Evidence from a Matching Model with Transfers (Job Market Paper, \href{https://github.com/MiriamLK/repo/raw/master/paper/Paper_2020_April_v1.pdf}{\underline{link}}) }.
\begin{itemize}
\item {I examine whether women remain concentrated in certain occupations because workers prefer to work with their own gender. I build a structural model in which workers maximize utility over occupation, wage, and the gender composition of their occupation, and firms maximize profit over the gender and wages of their employees. Using a Bartik instrumental variables strategy, I find that women strongly prefer to enter female-dominated occupations, but men show no evidence of gender preference. Given these estimates, equilibrium simulations indicate that equal pay for equal work laws could in fact increase segregation by preventing employers from compensating for gender preference.}
\end{itemize}
%\item { Encouraging Adoption of Urea Deep Placement Fertilizer in Liberia} (with Markus Goldstein and Niklas Buehren).
%\item{ When Competition Corrupts: A Theoretical Analysis of Market Structure and the Incidence of Corruption} (with Kaushik Basu and Boyang Zhang). World Bank Working Paper 6596.
%
%\item { The Algebra of Shared Prosperity and Some Empirical Regularities} (with David Rosenblatt). World Bank Working Paper 6645.
%
%\item { Child Brides, Bargaining Power, and Reform of Ethiopia's Family Law} (Job Market Paper).
\item { Intrahousehold Interaction and Married Family Labor Supply. \href{https://drive.google.com/file/d/1yG-O6b6oU1yLJThSJ9aXiIwwC4Hgt8Dr/view}{\underline{link}} }
\begin{itemize}
\item {Little is known about the role of household decision-making in the dramatic increase in labor supply of married women in the U.S. over the past half-century. In this paper, I use a bivariate probit and a Nash best-response game to look for changes in how married households supply labor over time in Census and PSID data. Consistent with past literature, I find in general that husbands prefer a household where one spouse specializes in home production and the other in market work, while wives prefer either both spouses work or both stay at home. Identification of these asymmetric effects comes from exclusion restrictions and sufficient variation in the characteristics of husbands and wives that push them to work in the market or at home. Next, I allow the interaction between spouses to depend on whether there are children in the home. I find that children increase complementarity of labor supply decisions for men and substitutability for women, but this has reversed over time for men. Lastly, I look at variation by spouse's labor income. Higher income husbands decrease the return to working for wives, but higher earning wives increase the return for husbands, consistent with a preference against being out-earned.}
\end{itemize}
\end{itemize}

%\subsection*{Work in Progress}
%\begin{itemize}
%\item { Intrahousehold Interaction and Married Family Labor Supply.}
%\end{itemize}

\section*{\scshape Grants, Awards, and Fellowships}
\begin{itemize} 
% \item Affiliate, Cornell Population Center, Cornell University, 2016.
  \item Sage Fellowship, Cornell University, 2012-2013.
  \item Virginia Galbraith Graduate Fellowship, Mount Holyoke College College Department of Economics, 2012.
  \item High Honor in Economics, for thesis:

  \hfill{The Effect of Child Care Subsidies on the Employment of Mothers, Mount Holyoke, 2012.}
  \item Phi Beta Kappa, Mount Holyoke College, 2012.
  %\item Sarah Williston Prize Winner, Mount Holyoke College, 2010.
 % \item Scholar Athlete Award, 2012.
  %\item Academic All-Conference Team (Track and Field), New England Women's and Men's Athletic Conference, 2010, 2012.
\end{itemize}

\section*{\scshape Teaching}
\begin{itemize} 
\itemsep0em
  \item TA, Labor Economics (Cornell ILRE 2400) for Prof. George Jakubson \hfill Spring 2018
  \item TA, Labor Economics (Cornell ILRE 2400 and 5400) for Prof. Gary Fields \hfill Fall 2017
  \item TA, Econometrics (Cornell Econ 3140) for Prof. Francesca Molinari \hfill Spring 2016
  \item TA, Intro to Econometrics (Mount Holyoke Econ 220) for Prof. Sarah Adelman \hfill Spring 2012
   \item Tutor, Economics Department (Mount Holyoke College) \hfill Fall 2011
\end{itemize}

%\section*{\scshape Invited Presentations and Workshops \small{(includes scheduled)}}
%\begin{itemize} 
%    \item Wellesley College, November 2015.
%    \item Cornell University, November 2015.
%    \item NBER Summer Institute, Labor Studies Workshop, invited participant, July 2015.
%    \item Committee on the Status of Women in the Economics Profession (CSWEP) session, Eastern Economics Association Annual Meeting, February 2015.
%    \item Advanced Graduate Workshop on Poverty, Development, and Globalization, Azim Premji University (with Columbia University), January 2015. 
% %   \item Phoenix Country Day School, Phoenix, AZ, November 2014.
%   \end{itemize}

\section*{\scshape Other Work Experience}

\begin{itemize}  
  \item {\it RA} for Prof. Mallika Thomas, Labor Economics, Cornell University ILR \hfill 2016-2017
  \item {\it RA} for Prof. Richard Mansfield, Labor Economics, Cornell University ILR \hfill Fall 2015
    \item {\it RA} for Prof. Lawrence Kahn, Labor Economics, Cornell University ILR \hfill  2014-2015
      \item {\it RA} for Prof. Levon Barseghyan, Economics, Cornell University \hfill Spring 2014
        \item {\it RA} for Prof. Kevin Hallock, Labor Economics, Cornell University ILR \hfill  Fall 2013
  \item {\it Junior Fellow Intern} National Center for Education Statistics, Washington D.C. \hfill Summer 2011
  \item {\it Undergraduate Researcher}, NSF-REU in mathematics at Central Michigan University \hfill Summer 2010 
  \item {\it Global Affairs Officer} Department of State, U.S. Embassy, Berlin, Germany \hfill Summer 2009
\end{itemize}

\section*{\scshape Professional Service}
\begin{itemize} 
\item 
\item {\it Refereed manuscripts for}, China Economic Review, Southern Economic Journal
\item {\it Chair}, \href{https://www.ftc.gov/news-events/events-calendar/twelfth-annual-federal-trade-commission-microeconomics-conference}{Panel: Learning about Substitution and Welfare from Data}, FTC Microeconomics Conf., 2019
\item {\it Co-coordinator}, \href{https://www.ftc.gov/about-ftc/bureaus-offices/bureau-economics/conferences-seminars/bureau-economics-seminar-series}{Bureau of Economics Seminar Series}, FTC, 2019-Present
\item {\it Co-organizer}, Labor Work in Progress Seminar, Cornell University, 2016-2017.
\item {\it Mentor}, Graduate Student Association for Economics Mentorship Program, Cornell University, 2015-2017
\end{itemize}

%
%\section*{\scshape Professional Affiliations}
%\begin{itemize} 
%\item American Economic Association.
%\item Eastern Economic Association.
%\end{itemize}

%
%\section*{\scshape Professional Experience }
%\begin{itemize}
% \setlength{\itemsep}{1pt}
 % \setlength{\parskip}{0pt}
 % \setlength{\parsep}{0pt}

%\item {\it Referee}, Economics Bulletin.
%\item {\it Discussant}, Eastern Economics Association meetings, February 2015.
%\item {\it President}, Graduate Student Association for Economics, Cornell University, 2012-2013.
% \item {\it Consultant}, World Bank Group, Office of the Chief Economist, Washington, DC  \hfill Summers 2013 and 2015
%\item {\it Senior Analyst}, Analysis Group, Boston, MA \hfill July 2008 -- February 2011
%\end{itemize}

% \begin{itemize}
% \item \textit{Programming Languages:} Fortran, C, Perl, Python, Emacs Lisp, \LaTeX, Stata, Mata, SQL (when under coercion, C++ and Matlab).
% \item \textit{Security Clearance:} Special Sworn Status, U.S. Census Bureau.
% \end{itemize}

%\section*{\scshape References}
%

%\begin{itemize}
%\item Professor \hfill {Phone:  607-255-6367}\\
%Department of Economics \hfill{E-mail: fm72@cornell.edu}\\
%Cornell University
%\end{itemize}
%
%%\halfblankline
%
%Richard Mansfield
%\begin{itemize}
%\item Assistant Professor\hfill {Phone:  303-735-5500} \\
%Department of Economics \hfill{E-mail: richard.mansfield@colorado.edu}\\
%University of Colorado Boulder
%\end{itemize}
%
%%\halfblankline
%
%%\halfblankline
%
%Victoria Prowse
%\begin{itemize}
%\item Magner Chair \hfill {Phone: 765-496-2049} \\ % cell \hfill {Phone:  607-227-3126}\\
%Associate Professor of Economics \hfill{E-mail: vprowse@purdue.edu}\\
%Purdue University
%\end{itemize}

%Mallika Thomas
%\begin{itemize}
%\item Assistant Professor \hfill {Phone: 607-255-5345}\\
%Department of Economics \hfill{E-mail: mallika.thomas@cornell.edu}\\
%Cornell University
%\end{itemize}

%\section*{\scshape Other}
%\begin{itemize}
% \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%%\item Leadership: {\it President}, Graduate Student Association for Economics, Cornell University, 2012-2013.
%\item Programing: Stata, Matlab, \LaTeX, Beamer, Excel
%\item Languages: English (native), German (proficient), French (proficient).
%\item Citizenship: USA.
%%\item Date of Birth: July 20, 1990
%%\item Extensive solo travel experience: India, China, Southeast Asia.
%\end{itemize}

% Footer
\medskip
\begin{center}
  \begin{small}
    Last updated: \today
  \end{small}
\end{center}

\end{document}